### 2018-01-04 submission/manuscript.Rmd
re-write last two paragraphs of results

Conversely to the stool models, both genera and OTU based models in tissue had the vast majority of their top 10 occur in a study specific manner [Figure S7]. *Fusobacterium* and *Fusobacteriaceae* show up more often in the top 10 for matched tissue samples but were not present in the top 10 or were much lower ranked for unmatched tissue [Figure S7B-C & S7E-F]. There appears to be very little overlap in the top 10 most important variables between stool and tissue for both adenoma and carcinoma [Figure 6 & S7].

***A Majority of Studies are Underpowered for Detecting Small Effect Size Differences:***
Based on the previous observations we then assessed whether the studies that we included are realistically powered to identify small, medium, and large scale differences between case and control. When assessing the power of each study at different effect sizes the majority of studies for both adenoma and carcinoma have an 80% power to detect a 30% difference [Figure 7A & B]. No single study that was analyzed had the standard 80% power to detect an effect size difference that was equal to or below 10% [Figure 7A & B].  In order to achieve adequate power for small effect sizes, studies would need to recruit over 1000 individuals for each arm [Figure 7C].

>>>>

Unlike stool, tissue based Random Forest models showed no consistent representation of *Ruminococcaceae*, *Ruminococcus*, *Bacteroides*, and *Roseburia* in the top 10 across study [Figure S7]. The vast majority of the top 10 genera and OTUs occured in a study specific manner for these tissue based Random Forest models. For both the genera and OTU Random Forest models of adenoma and carcinoma, there appears to be very little overlap in the top 10 most important variables between stool and tissue [Figure 6 & S7]. This discordance between stool and tissue also applies to the mouth-associated genera with one noticeable skew being that *Fusobacterium* and *Fusobacteriaceae* occur more often in the top 10 of Random Forest models using matched versus unmatched tissue samples [Figure S7B-C & S7E-F].

***CRC Studies are Underpowered for Detecting Small Effect Sizes:***
Based on our reported results we assessed whether the studies analyzed were realistically powered to identify small, medium, and large scale differences between cases and controls. When assessing the power of each study at different effect sizes the majority of studies, for both adenoma and carcinoma, acheived 80% power to detect a 30% or greater difference between groups [Figure 7A & B]. No study that was analyzed had the standard 80% power to detect an effect size difference that was equal to or below 10% [Figure 7A & B].  In order to achieve a power of 80%, for small effect sizes, studies used in our meta-analysis would need to recruit over 1000 individuals for each arm [Figure 7C].

### 2018-01-04 submission/manuscript.tex
re-write last two paragraphs of results

Conversely to the stool models, both genera and OTU based models in
tissue had the vast majority of their top 10 occur in a study specific
manner {[}Figure S7{]}. \emph{Fusobacterium} and \emph{Fusobacteriaceae}
show up more often in the top 10 for matched tissue samples but were not
present in the top 10 or were much lower ranked for unmatched tissue
{[}Figure S7B-C \& S7E-F{]}. There appears to be very little overlap in
the top 10 most important variables between stool and tissue for both
adenoma and carcinoma {[}Figure 6 \& S7{]}.

\textbf{\emph{A Majority of Studies are Underpowered for Detecting Small
Effect Size Differences:}} Based on the previous observations we then
assessed whether the studies that we included are realistically powered
to identify small, medium, and large scale differences between case and
control. When assessing the power of each study at different effect
sizes the majority of studies for both adenoma and carcinoma have an
80\% power to detect a 30\% difference {[}Figure 7A \& B{]}. No single
study that was analyzed had the standard 80\% power to detect an effect
size difference that was equal to or below 10\% {[}Figure 7A \& B{]}. In
order to achieve adequate power for small effect sizes, studies would
need to recruit over 1000 individuals for each arm {[}Figure 7C{]}.

>>>>

Unlike stool, tissue based Random Forest models showed no consistent
representation of \emph{Ruminococcaceae}, \emph{Ruminococcus},
\emph{Bacteroides}, and \emph{Roseburia} in the top 10 across study
{[}Figure S7{]}. The vast majority of the top 10 genera and OTUs occured
in a study specific manner for these tissue based Random Forest models.
For both the genera and OTU Random Forest models of adenoma and
carcinoma, there appears to be very little overlap in the top 10 most
important variables between stool and tissue {[}Figure 6 \& S7{]}. This
discordance between stool and tissue also applies to the
mouth-associated genera with one noticeable skew being that
\emph{Fusobacterium} and \emph{Fusobacteriaceae} occur more often in the
top 10 of Random Forest models using matched versus unmatched tissue
samples {[}Figure S7B-C \& S7E-F{]}.

\textbf{\emph{CRC Studies are Underpowered for Detecting Small Effect
Sizes:}} Based on our reported results we assessed whether the studies
analyzed were realistically powered to identify small, medium, and large
scale differences between cases and controls. When assessing the power
of each study at different effect sizes the majority of studies, for
both adenoma and carcinoma, acheived 80\% power to detect a 30\% or
greater difference between groups {[}Figure 7A \& B{]}. No study that
was analyzed had the standard 80\% power to detect an effect size
difference that was equal to or below 10\% {[}Figure 7A \& B{]}. In
order to achieve a power of 80\%, for small effect sizes, studies used
in our meta-analysis would need to recruit over 1000 individuals for
each arm {[}Figure 7C{]}.
